\addchap{\lsAcknowledgementTitle} 

This book was written, with interruptions, over a period of 15 years, during which I used many versions of it in my seminars and communicated with dozens of colleagues about methodological issues I came up against in writing it. My thanks go to my students at the
Universität Bremen, the Universität Hamburg and the Freie Universität Berlin and to Harald Baayen, Claudia Baldermann, Michael Barlow, Timothy Colleman, Alice Deignan, Holger Diessel, Stefan Evert, Kerstin Fischer, Susanne Flach, Adele Goldberg, Juliana Goschler, Stefan Gries, Beate Hampe, Stefan Hartmann, Thomas Herbst, Jürgen Hermes, Martin Hilpert, Veronika Koller, Jeong-Hwa Lee, Yoo Yung Lee, Britta Mondorf, Carlos Nash, Erich Neuwirth, Klaus-Uwe Panther, Günter Radden, Alexander Rauhut, Ada Rohde, Doris Schönefeld, Peter Uhrig, Daniel Wiechmann, Stefanie Wulff, Hilary Young, Arne Zeschel, and to everyone I have forgotten here. A few of these people deserve a more detailed mention for the substantial roles they played in the emergence of this book. Klaus-Uwe Panther and Günter Radden encouraged my first steps into corpus linguistics when supervising my master's thesis, and Michael Barlow continued this encouragement during my time as a Ph.D. student at Rice University. My fellow student Ada Rohde and I spent many hours sifting through concordances for our dissertations and discussing questions of retrieval and annotation over a beer or two. My collaboration with Stefan Gries during the first decade of the new millennium was an exciting and most instructive time in my research career. Susanne Flach read and commented on every chapter of the final version of the book, which I started preparing in 2014, and she also introduced me to the excellent Corpus Workbench environment, without which the book would have taken another 15 years. Yoo Yung Lee proof-read the manuscript and checked many of the queries. Two anonymous reviewers, one of whom revealed himself to be Stefan Hartmann, provided stern but constructive comments on the manuscript. Stefan Müller's founding of the open-access Language Science Press provided me with an incentive to finally get serious about finishing this book, and Felix Kopecky and Sebastian Nordhoff always remained friendly and supportive when {\LaTeX} drove me to dispair. Finally, Juliana Goschler has been, and continues to be, a wonderful colleague and the light of my life.
